### 2015-02-21 writing/F'14 paper/discussion.tex
Paper presentation tweaks; SNLI setup; misc updates

Of course, challenges remain. Even
the RNTN falls short of perfection in the recursion experiment, with
performance falling off steadily as the size of the expressions grows. It
remains to be seen whether these deficiencies can be overcome with
stronger models or optimization procedures. The space of possible stronger
models is sizable, including extensions of the recursive models used here
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, recurrent
models \cite{sutskever2014sequence}, and even learning-enabled variants 
of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
In addition,
there remain subtle questions about how to fairly assess whether these
models have truly generalized in the way we want them to. There is a
constant tension between showing the models training data that gives
them a chance to learn the target logical functions and revealing the
answer to them in a way that leads to overfitting. The underlying
logical theories provide only limited guidance on this point.

Our artificial data experiments have only scratched the surface 
of the logical complexity of natural language; in future experiments, we hope to test sentences
with embedded quantifiers, multiple interacting quantifiers, relative
clauses, and other kinds of recursive structure. Our SICK experiments 
similarly only begin to reveal the potential of these models to learn to 
perform complex semantic inferences from corpora. Nonetheless, the
rapid progress the field has made with these models in recent years
provides ample reason to be optimistic that they can be trained to
meet all the challenges of natural language semantics.

>>>>

Of course, challenges remain. Even
the RNTN falls short of perfection in the recursion experiment, with
performance falling off steadily as the size of the expressions grows. It
remains to be seen whether these deficiencies can be overcome with
stronger models or optimization procedures. The space of possible stronger
models is sizable, including extensions of the recursive models used here
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, recurrent
models \cite{sutskever2014sequence}, and even learning-enabled variants 
of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
In addition, interesting analytical questions remain about \ii{how} these models encode
the underlying logics. Neither the underlying
logical theories, nor any straightforward parameter inspection technique provides 
much insight on this point, but we hope that further experiments may reveal 
some structure in the representations.

Our artificial data experiments have only scratched the surface 
of the logical complexity of natural language; in future experiments, we hope to test sentences
with more complex types of structure, as well as run similar experiments on non-tree structured models, such as LSTMs. Our SICK experiments 
similarly only begin to reveal the potential of these models to learn to 
perform complex semantic inferences from corpora. Nonetheless, the
rapid progress the field has made with these models in recent years
provides ample reason to be optimistic that they can be trained to
meet all the challenges of natural language semantics.

### 2015-02-21 writing/F'14 paper/intro.tex
Paper presentation tweaks; SNLI setup; misc updates

Supervised recursive neural network models (RNNs) for sentence meaning
have been successful in an array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,irsoydeep},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
the ability of these models to learn compositional semantic grammars.
However, it remains an open question whether they can achieve the
high-fidelity distributed representations of recent algebraic
approaches \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
and whether they can match the performance of grammars based in logical forms
\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
on the core semantic issues of quantification,
entailment, and contradiction needed to identify the relationships
between sentences like \ii{every animal walks}, \ii{every turtle
 moves}, and \ii{most reptiles don't move}.

@@

Our research uses the task of \ii{natural language inference} 
(also known as \ii{recognizing textual entailment};
\citealt{dagan2006pascal}), in which the goal is to determine the core
inferential relationship between two sentences. Much of the
theoretical work on this task (and some successful implemented models;
\citealt{maccartney2009extended,watanabe2012latent}) involves \ii{natural
  logics}, which are formal systems that define rules of inference
between natural language words, phrases, and sentences without the
need of intermediate representations in an artificial logical
language. We use the natural logic developed by
\newcite{maccartney2009extended} as our formal model. This logic
defines seven core relations of synonymy, entailment, contradiction,
and mutual consistency, as summarized in Table~\ref{b-table}, and it
provides rules of semantic combination for projecting these relations
from the lexicon up to complex phrases. The formal properties and
inferential strength of this system are now well-understood
\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.

>>>>

% TODO: TreeRNN?
% TODO: More language about the representations being the object of study.

Supervised recursive neural network models (RNNs, \newcite{goller1996learning}) for sentence meaning
have been successful in an array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi,irsoydeep},
image description \cite{sochergrounded}, and paraphrase detection
\cite{Socher-etal:2011:Paraphrase}. These results are encouraging for
the ability of these models to learn compositional semantic grammars.
However, it remains an open question whether they can achieve the
high-fidelity distributed representations of recent algebraic
approaches \cite{ClarkCoeckeSadrzadeh2011,grefenstette2013towards,Hermann-etal:2013,rocktaschellow},
and whether they can match the performance of grammars based in logical forms
\cite{Warren:Pereira:1982,Zelle:Mooney:1996,ZetCol:2005,LiangJordan:2013}
on the core semantic issues of quantification,
entailment, and contradiction needed to identify the relationships
between sentences like \ii{every animal walks}, \ii{every turtle
 moves}, and \ii{most reptiles don't move}.

@@

Our research uses the task of \ii{natural language inference} 
(also known as \ii{recognizing textual entailment};
\citealt{dagan2006pascal}), in which the goal is to determine the core
inferential relationship between two sentences. Much of the
theoretical work on this task (and some successful implemented models;
\citealt{maccartney2009extended,watanabe2012latent}) involves \ii{natural
  logics}, which are formal systems that define rules of inference
between natural language words, phrases, and sentences without the
need of intermediate representations in an artificial logical
language. We use the natural logic developed by
\newcite{maccartney2009extended} as our formal model. This logic
defines seven mutually-exclusive relations of synonymy, entailment, contradiction,
and mutual consistency, as summarized in Table~\ref{b-table}, and it
provides rules of semantic combination for projecting these relations
from the lexicon up to complex phrases. The formal properties and
inferential strength of this system are now well-understood
\cite{Icard:Moss:2013,Icard:Moss:2013:LILT}.

### 2015-02-21 writing/F'14 paper/join.tex
Paper presentation tweaks; SNLI setup; misc updates

We trained versions of both the RNN model and the RNTN model on these
data sets. In both cases, the models are implemented as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer is not involved, and the two models differ only in which
layer function is used for the comparison layer. We simply present
the models with the (randomly initialized) embedding vectors for each
of two terms, ensuring that the model has no information about the terms
being compared except for the relations between them that appear in training.

@@

      $a \nateq b$              & $b \nateq b$ \\
      $a \natfor e$              & $b \natfor e$ \\
      $a \natneg g$              & \strikeout{$e \nateq f$} \\
      $d \natfor h$              & \strikeout{$g \natrev d$} \\
      $e \natalt g$            & $h \natrev d$ \\

>>>>

We trained versions of both the RNN model and the RNTN model on these
data sets\footnote{Since this task relies entirely on the learning of the two entity vectors, no obvious comparable baseline presents itself.}. In both cases, the models are implemented as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer is not involved, and the two models are not recursive and differ only in which
layer function is used for the comparison layer. We simply present
the models with the (randomly initialized) embedding vectors for each
of two terms, ensuring that the model has no information about the terms
being compared except for the relations between them that appear in training.

@@

      $a \nateq b$              & $b \natneg g$ \\
      $a \natrev e$              & $b \natrev e$ \\
      $d \natrev h$              & \strikeout{$e \nateq f$} \\
       $e \natalt g$              & \strikeout{$g \natfor d$} \\
      $g \natneg a$           & $h \natfor d$ \\

### 2015-02-21 writing/F'14 paper/quantifiers.tex
Paper presentation tweaks; SNLI setup; misc updates

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{25d RNN}  & \multicolumn{2}{c}{25d RNTN}\\
    \midrule
    Train & 35.4 &(7.5) &	\textbf{100}&		\textbf{(100)}&	99.5&	(99.3)\\	
    Test & 35.4 &(7.5) &	\textbf{99.4}&\textbf{(99.5)}& \textbf{99.4} & (99.2)\\
    \bottomrule
  \end{tabular}
  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}
  \label{qresultstable}
\end{table} 
%
% do not allow a blank line --- adds too much space
%
\paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic satisfactorily. We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

>>>>

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{25d RNN}  & \multicolumn{2}{c}{25d RNTN}\\
    \midrule
    Train & 35.4 &(7.5) &	\textbf{100}&		\textbf{(100)}&	99.5&	(99.3)\\	
    Test & 35.4 &(7.5) &	\textbf{99.4}&\textbf{(99.5)}& \textbf{99.4} & (99.2)\\
    \bottomrule
  \end{tabular}
  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}
  \label{qresultstable}
\end{table} 
%
% do not allow a blank line --- adds too much space
%
\paragraph{Results} The results, reported in Table~\ref{qresultstable}, show that both models are able to learn to generalize the underlying logic satisfactorily. We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

% TODO: Baseline

### 2015-02-21 writing/F'14 paper/recursion.tex
Paper presentation tweaks; SNLI setup; misc updates

\begin{table}[tp]
  \centering\small
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{l l}
      \toprule
      Formula     & Interpretation \\
      \midrule
      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
      \bottomrule
    \end{tabular}    
    \caption{Well-formed formulae. $\varphi$ and $\psi$
      range over all well-formed formulae, and $\sem{\cdot}$ is
      the interpretation function mapping formulae into $\{\True,
      \False\}$.}\label{tab:pl}
  \end{subtable}
  \begin{subtable}[t]{0.45\textwidth}
    \centering\vspace{4mm}
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg a$        & $\natneg$ & $a$ \\
      $\plneg \plneg a$ & $\nateq$  & $a$ \\
      $a$               & $\natfor$ & $(a \plor b)$ \\
      $a$               & $\natrev$ & $(a \pland b)$ \\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg\, (\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
      \bottomrule
    \end{tabular}
    \caption{Examples of statements about relations between
      well-formed formulae, defined in terms of sets of satisfying
      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
  \end{subtable}
  \caption{Natural logic relations over sentences of propositional logic.}  
  \label{prop-figure}
\end{table}

Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector RNN
model somewhat similar to our RNTN can learn boolean logic, 
a logic where the atomic symbols are simply the
values $0$ and $1$. While learning the operators of that logic is not trivial, the outputs of
each operator can be represented accurately by a single bit.
In the propositional logic used here, the atomic symbols are variables over those values. 
Statements of this logic thus describe conditions on
unknown truth values of propositions. This logic distinguishes
between $2^{6} = 64$ possible assignments of truth values, and
expressions of this logic define arbitrary conditions on these
possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
possible statements that the intermediate vector representations need
to be able to distinguish if we are to learn this logical system in full.

In our experiments, we randomly generate a large set of unique pairs 
of formulae and compute the relation that holds for each pair.
We discard pairs in which either statement is a tautology or
contradiction (a statement that is true of either all or no possible
assignments), for which none of the seven relation labels in
Table~\ref{b-table} can hold. The resulting set of formula pairs is
then partitioned into bins based on the number of logical operators in
the longer of the two formulae. We then randomly sample 20\% of each
bin for a held-out test set.

@@

@@

\paragraph{Results} 
We trained both the RNN and RNTN models on data with up to 4
connectives (60k pairs) and tested it on a separate set of examples of up to 12
connectives (21k pairs). We initialized the model parameters randomly,
including the vector representations of the six variables.
Figure~\ref{prop-results} shows test the relationship
between test accuracy and statement length. We found that both 
models were able to perform well on unseen small test examples, 
with RNN accuracy above
98\% and RNTN accuracy above 99\% on formulae below length five.
Starting at size four, performance gradually falls with increasing
size. Training accuracy was 99.4\% for the RNN and 99.8\% for the RNTN.

>>>>

\begin{table}[tp]
  \centering\small
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{l l}
      \toprule
      Formula     & Interpretation \\
      \midrule
      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
      \bottomrule
    \end{tabular}    
    \caption{Well-formed formulae. $\varphi$ and $\psi$
      range over all well-formed formulae, and $\sem{\cdot}$ is
      the interpretation function mapping formulae into $\{\True,
      \False\}$.}\label{tab:pl}
  \end{subtable}
  \begin{subtable}[t]{0.45\textwidth}
    \centering\vspace{4mm}
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg a$        & $\natneg$ & $a$ \\
      $\plneg \plneg a$ & $\nateq$  & $a$ \\
      $a$               & $\natfor$ & $(a \plor b)$ \\
      $(a \plor (b \plor c))$               & $\natrev$ & $(b \pland  \plneg c)$ \\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg\, (\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
      \bottomrule
    \end{tabular}
    \caption{Examples of the type of statements used for training and testing. These are relations between
      well-formed formulae, computed in terms of sets of satisfying
      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
  \end{subtable}
  \caption{Natural logic relations over sentences of propositional logic.}  
  \label{prop-figure}
\end{table}

Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector RNN
model somewhat similar to our RNTN can learn boolean logic, 
a logic where the atomic symbols are simply the
values $0$ and $1$. While learning the operators of that logic is not trivial, the outputs of
each operator can be represented accurately by a single bit.
In the more demainding task presented here, the atomic symbols are variables over binary values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions over valuations of 
these variables. Relational expressions between these conditions are essentially theorems of propositional logic, and to succeed, the model must learn to approximate the behavior of a theorem prover.

In our experiments, we randomly generate unique pairs 
of formulae and compute the relation that holds for each pair.
We discard pairs in which either statement is either a tautology or a
contradiction, for which the seven relations in
Table~\ref{b-table} are undefined. The resulting set of formula pairs is
then partitioned into 13 bins according the number of logical operators in
the longer of the two formulae. We then sample 20\% of each
bin for a held-out test set.

@@

In order to test the model's generalization to unseen structures, we discard
training examples with more than 4 connectives, yielding 60k short training examples,
and 21k training examples containing up to 12 connectives.
We evaluated both the RNN and RNTN\footnote{We did not run any strictly simpler baselines for this experiment because it is demonstrably impossible to perform substantially above chance on longer expressions without word order information.} on the resulting data, initializing the model parameters randomly,
including the vector representations of the six variables.

@@

\paragraph{Results} Figure~\ref{prop-results} shows test the relationship
between test accuracy and statement length. We found that both 
models were able to perform well on unseen small test examples, 
with RNN accuracy above
98\% and RNTN accuracy above 99\% on formulae below length five.
Starting at size four, performance gradually falls with increasing
size. Training accuracy was 99.4\% for the RNN and 99.8\% for the RNTN.

### 2015-02-21 writing/F'14 paper/sick.tex
Paper presentation tweaks; SNLI setup; misc updates

\begin{table*}[htp]
  \centering\small
  \begin{tabular}{lcl}
    \toprule
A woman is not boiling shrimps.& \ii{contradiction}&	A woman is boiling shrimps.\\
A wild deer is jumping a fence. &\ii{entailment}	&A deer is jumping a fence.\\
A potato is being sliced by a woman. &\ii{entailment}	&A woman is slicing a potato.\\
The plane, which is south African, is flying in a blue sky.& \ii{entailment}&	An airplane is flying through the air.\\
    \bottomrule
  \end{tabular}
  \caption{\label{examplesofsickdata}Examples of correct RNTN classifications on SICK.}
\end{table*}

@@

We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set. We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}), a corpus of automatically labeled (and thus noisy) entailment examples over image captions, the same genre of text from which SICK was drawn. We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source. We parsed the data from both sources with the Stanford PCFG Parser v. 3.3.1 \cite{klein2003accurate}. We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word. The training and test data on which we report performance are collapsed in this way, and both collapsed and uncollapsed copies of the training data are used in training. Finally, in order to improve regularization on the noisier data, we used dropout \cite{srivastava2014dropout} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).

@@

>>>>

\begin{table*}[htp]
  \centering\small
  \begin{tabular}{lcl}
    \toprule
A woman is not boiling shrimps.& \ii{contradiction}&	A woman is boiling shrimps.\\
A wild deer is jumping a fence. &\ii{entailment}	&A deer is jumping a fence.\\
A potato is being sliced by a woman. &\ii{entailment}	&A woman is slicing a potato.\\
The plane, which is south African, is flying in a blue sky.& \ii{entailment}&	An airplane is flying through the air.\\
    \bottomrule
  \end{tabular}
  \caption{\label{examplesofsickdata}Examples of correct RNTN classifications on SICK.}
\end{table*} % TODO: Update. Find incorrect examples?

@@

We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set. We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}, also used by the winning SICK submission), a corpus of noisy automatically labeled entailment examples over image captions, the same genre of text from which SICK was drawn. We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source. We parsed the data from both sources with the Stanford PCFG Parser v. 3.3.1 \cite{klein2003accurate}. We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair down to a single head word. The training and test data on which we report performance are collapsed in this way, and both collapsed and uncollapsed copies of the training data are used in training. Finally, in order to improve regularization on the noisier data, we used dropout \cite{srivastava2014dropout} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).

@@

% TODO: Update results
